Using pressure-driven flow systems to evaluate laser speckle contrast imaging

### 2022-12-12 article.tex
Add T32 to funding
Initial draft of reviewer response

Due to its popularity, LSCI instrument development is an active area of research with microfluidic studies commonly used to characterize performance in controlled flow environments. Typically, studies designed to mimic blood flow involve the use of a flow channel embedded in a polydimethylsiloxane (PDMS) substrate \cite{Parthasarathy:2008el}, glass capillary tubing \cite{Arias-Cruz.2019}, or clear plastic tubing \cite{Miao:2014}. More complex microfluidic networks have also been fabricated in glass, plastic, or epoxy substrates to represent superficial heterogenous vasculature \cite{Luu.2012}. A scattering solution \cite{Parthasarathy:2008el,Arias-Cruz.2019} or whole blood \cite{Miao:2014} is then flowed through the channel at controlled rates and imaged using LSCI. For applications in rodent studies, 0.1 -- 10 mm/s is a commonly selected range of flow speeds that corresponds with experimentally measured capillary speeds \cite{Tomita:2008do}. This speed can be converted to a volumetric flow rate using the cross-sectional area of the flow channel, which is then used to program the flow control system.

Syringe pumps are the most widely used flow control system in microfluidic studies and are believed to provide a highly-controlled flow output. A survey of over twenty LSCI studies using microfluidics that were published since 2015 found that syringe pumps were used in the majority (90\%) of experiments. The devices were from a variety of manufacturers and featured various syringe sizes and materials. Syringe pumps operate by applying mechanical force on the plunger of a syringe at a controlled speed. A rotating lead screw driven by a stepper motor controls the velocity of the plunger, which is calculated based on the syringe size and inner diameter to achieve a desired flow rate. However, syringe pumps produce oscillations in flow with varying frequency and magnitude due to the physical motion of the motor and from imperfections on the lead screw \cite{Li:2014ca,Zeng.2014}. Syringe pumps have also been shown to have slow responsivity and can take a long time to achieve programmed flow rates \cite{Zhou:2011ey}. The trade-off between stability and responsivity makes it difficult \cite{Kang.2012} to obtain both stable flow and a fast flow response with a syringe pump system, both of which are necessary for LSCI characterization measurements.

Pressure-driven flow control systems overcome the limitations of syringe pumps by eliminating the use of a motor and lead screw to provide very stable flow \cite{Zhou:2011ey,Li:2014ca,Zeng.2020}. These systems use pressure-regulated air to push liquid from a reservoir through the microfluidic device at a constant flow rate \cite{Zeng.2020}. In order to control the applied pressure, an absolute flow sensor must be added inline with the microfluidic device. Thermal mass flow sensors are commonly used to measure the low flow rates found in microfluidics. These sensors use heating and temperature-sensing elements to detect the magnitude and direction of flow and are factory-calibrated to measure absolute flow rates for various fluids. Integration of the flow sensor with a feedback loop allows for the programming of specific flow rates rather than pressure levels.

@@

The microfluidic flow phantom was cast using PDMS (Sylgard 184™, Dow Corning) with 1.8 mg titanium dioxide (\ce{TiO2}) added per gram of PDMS to generate a scattering background that mimics tissue optical properties at near infrared wavelengths ($\mu_s' = 8$ cm$^{-1}$) \cite{Parthasarathy:2008el,Yaroslavsky:2002tg}. The PDMS \ce{TiO2} mixture was vigorously mixed for 15 minutes to ensure even distribution of the \ce{TiO2} in the phantom. Centrifugation (3000 g relative centrifugal force, 10 minutes) was then performed to de-gas the sample and remove any \ce{TiO2} clumps.

A machined aluminum mold was used to cast the microfluidic as shown in Fig.~\ref{fig:microfluidic}a. The extrusion is 460 $\mu$m $\times$ 460 $\mu$m $\times$ 25 mm, which formed three walls of the channel. The square cross-section (1:1 aspect ratio) approximates the circular cross-section of vascular geometries \textit{in vivo}. The PDMS \ce{TiO2} mixture was carefully poured on top of the mold within a polystyrene petri dish and placed in a vacuum chamber to remove air bubbles introduced during pouring. The PDMS was then cured at 70 $^\circ$C for one hour, allowed to cool, and detached from the mold. The microchannel side of the phantom was oxygen plasma bonded to a glass slide to seal the channel. Inlet and outlet ports were attached using 23 gauge blunt tip needles fitted partially inside the phantom, which were epoxied in place for an air-tight seal. Tygon® microbore tubing (0.02$''$ inner diameter, Cole-Parmer Instrument Company, LLC) was connected to the blunt tip ports as inlet and outlet tubing. The inlet tubing was connected to the flow control system using fluidic fittings with FEP tubing (IDEX Health and Science, LLC). The outlet tubing drained into a vial, with the tubing end fully submerged in liquid (see Supplementary Material S1). A photograph of the phantom is shown in Fig.~\ref{fig:microfluidic}b, where the channel is visible inside the white PDMS with the glass slide sealing the surface.

@@

To assess the performance of the syringe pump and pressure-regulated flow systems, and to quantify the uncertainty of each system, it is necessary to know the true flow rate of the blood-mimicking solution through the microfluidic channel. A thermal mass flow sensor (FLU-L, 0.03 -- 1 mL/min, Fluigent, Inc.) was placed inline between the flow control system and the phantom inlet tubing, as shown in Fig.~\ref{fig:system_lsci}a. The sensor output was digitized and relayed to a computer via an accompanying hub (Flowboard, FLB, Fluigent, Inc.) where the MAESFLO software (Fluigent, Inc.) was used to monitor and record the absolute flow rate. All flow sensor measurements were downsampled from 10 Hz to 2 Hz using interpolation to match the timing of the optical measurements.

@@

Two stepped flow profile experiments were conducted to assess each flow system using the commercial flow sensor. First, a 3-step flow profile was used to determine the uncertainty of each flow control system. The experiment lasted 240 seconds, beginning with 30 seconds of no flow (0 mm/s) followed by one full minute at each flow speed (2.4, 3.6, and 4.8 mm/s) before concluding with an additional 30 seconds of no flow. The experiment was repeated three times on each system. The second flow profile experiment consisted of a 13-step flow profile. The flow speed was once again varied from 2.4 to 4.8 mm/s, however in smaller 0.2 mm/s increments. As with the first experiment, the flow profile begins and ends with 30 seconds of no flow, with one full minute at each programmed flow speed (840 seconds total duration). The second experiment was also repeated three times on each flow control system.

@@

\noindent where $T$ is the camera exposure time, $x=T/\tau_{c}$, and $\beta$ is an instrumentation factor that accounts for
speckle sampling, polarization, and coherence effects \cite{Bandyopadhyay:2005bg}. For these measurements, it was assumed that $\beta=1$, which exclusively allows for the computation of relative flow changes. The resulting $\tau_c$ estimates were resampled to 2 Hz using interpolation and smoothed temporally with a central moving average filter ($k=5$).

@@

A pair of consumer camera lenses were used to image the scattered light (AF Nikkor 50 mm $f$/1.8D + AF Micro-Nikkor 105 mm $f$/2.8D, Nikon Corp.) with 2.1$\times$ magnification through a bandpass filter (785$\pm$31 nm, 87-773, Edmund Optics Inc.) to a CMOS camera (acA1920-155um, 1920 $\times$ 1200 pixels, Basler AG). The aperture of the lower lens was set to $f$/5.6 while the upper lens was left at its maximum size of $f$/2.8. Only a subset of the overall sensor array was used (1000 $\times$ 750 pixels), resulting in a field-of-view of 2.9 $\times$ 2.2 mm. The camera exposures were temporally synchronized with the modulated laser pulses. Fifteen different camera exposures ranging between 50 $\mu$s and 80 ms were recorded for each complete MESI frame, resulting in an effective acquisition rate of approximately 2.5 fps. The total amount of light used to capture each exposure was held constant with the AOM in order to minimize the effects of shot noise \cite{Parthasarathy:2008el}. The entire acquisition was controlled using custom software written in C++ integrated with a multifunction I/O device (USB-6363, National Instruments Corp.) for the generation of camera exposure trigger signals and AOM modulation voltages.

@@

A stepped flow profile experiment was conducted using the pressure-regulated flow system to assess the performance of both MESI and LSCI using the same optical hardware. The flow profile consisted of 19 steps spanning 1 -- 10 mm/s in 0.5 mm/s increments. Each speed was maintained for two minutes with a total duration of 40 minutes, including brief periods of no flow (0 mm/s) at the beginning and end of the profile. MESI measurements were acquired continuously throughout the experiment at the maximum acquisition rate of the system ($\sim$2.5 fps). Single-exposure (0.5 ms) LSCI frames were extracted from the MESI dataset \textit{post hoc}. The initial rise of the first flow step was again used to synchronize the timing of the MESI/LSCI and flow sensor datasets. The experiment was repeated a total of four times across two days.

@@

\noindent where $T$ is the camera exposure time, $x=T/\tau_c$, $\rho$ is the fraction of light that is dynamically scattered, $\beta$ is a normalization factor that accounts for speckle sampling, and $\nu$ represents exposure-independent instrument noise and nonergodic variances. $\beta$ was held constant at all timepoints during an experiment by performing an initial fit to Eq. \ref{eq:mesi} using the median speckle contrast value at each exposure across the entire multi-exposure dataset. All fitting was performed with the Levenberg-Marquardt nonlinear least squares algorithm \cite{Lourakis:J2fCMU5i} using a custom program written in MATLAB (R2021a, MathWorks, Inc.). The resulting $\tau_c$ estimates were resampled to 2 Hz using interpolation, smoothed temporally with a central moving average filter ($k=5$), and inverted to obtain $ICT$.

@@

Almost all microfluidic studies used for characterizing LSCI and MESI systems have been performed using syringe pumps. This study corroborates previous work \cite{Li:2014ca,Zhou:2011ey,Zeng.2020} that has shown syringe pumps do not produce stable flow outputs and demonstrates that single-exposure LSCI is sensitive enough to detect the resulting flow anomalies (Fig. \ref{fig:SvP-FS_LSCI}). Furthermore, the close agreement between LSCI and the flow sensor demonstrates that these anomalies could be incorrectly interpreted as noise in the optical imaging system when they are actual flow variations. Compared to the syringe pump, the pressure-regulated flow system produced much more stable and reproducible flow outputs across all trials. Because microfluidic studies are commonly used to assess LSCI and MESI instrument performance, an unstable flow control system contributes additional noise to the measurements, making it more difficult to properly characterize a system. For example, the flow profiles in Fig. \ref{fig:SvP-FS} and the uncertainty analysis in Table \ref{tab:uncertainty_combined} show large errors and high variability within and across trials using the syringe pump. These instabilities in flow were caused entirely by the syringe pump, as independently validated using single-exposure LSCI in Fig. \ref{fig:SvP-FS_LSCI}a. Such instability would increase the uncertainty in the true performance of any instrument being evaluated. The same analysis shows that a more reliable flow controller, such as the pressure-regulated flow system, produces minimal error and much improved reproducibility, therefore facilitating a more robust instrument assessment.

@@

The linearity of the results seen with both MESI and LSCI suggest that a calibration to absolute flow speeds might be feasible. However, any such calibration would be highly dependent upon experimental parameters, including the sample geometry (i.e. channel size), scatterer concentration, illumination configuration, and exposure time selection. Changes to any of these parameters would produce different $ICT$ values, which would affect the correlation with the flow sensor and invalidate the calibration. It also remains to be seen whether calibrations performed in controlled microfluidic environments can reliably be translated to animal or human subjects since \textit{in vivo} measurements sample a range of different vessels \cite{Davis:2016ik,Jafari.2020}.

@@

\subsection*{Acknowledgments}
This study was supported by the National Institutes of Health (Nos. NS108484 and EB011556) and the UT Austin Portugal Program.

>>>>

Due to its popularity, LSCI instrument development is an active area of research with microfluidic studies commonly used to characterize performance in controlled flow environments. Typically, studies designed to mimic blood flow involve the use of a flow channel embedded in a polydimethylsiloxane (PDMS) substrate \cite{Parthasarathy:2008el}, glass capillary tubing \cite{Arias-Cruz.2019}, or clear plastic tubing \cite{Miao:2014}. More complex microfluidic networks have also been fabricated in glass, plastic, or epoxy substrates to represent superficial heterogenous vasculature \cite{Luu.2012}. A scattering solution \cite{Parthasarathy:2008el,Arias-Cruz.2019} or whole blood \cite{Miao:2014} is then flowed through the channel at controlled rates and imaged using LSCI. For applications in rodent studies, 0.1 -- 10 mm/s is a commonly selected range of flow speeds that corresponds with experimentally measured capillary speeds \cite{Tomita:2008do}. This speed can be converted to a volumetric flow rate using the cross-sectional area of the flow channel, which can then be used to program the flow controller.

Syringe pumps are the most widely used flow control system in microfluidic studies and are generally believed to provide a highly-controlled flow output. A survey of over twenty LSCI studies using microfluidics that were published since 2015 found that syringe pumps were used in the majority (90\%) of experiments. The devices were from a variety of manufacturers and featured various syringe sizes and materials. Syringe pumps operate by applying mechanical force on the plunger of a syringe at a controlled speed. A rotating lead screw driven by a stepper motor controls the velocity of the plunger, which is calculated based on the syringe size and inner diameter to achieve a desired flow rate. However, syringe pumps produce oscillations in flow with varying frequency and magnitude due to the physical motion of the motor and from imperfections on the lead screw \cite{Li:2014ca,Zeng.2014}. Syringe pumps have also been shown to have slow responsivity and can take a long time to achieve programmed flow rates \cite{Zhou:2011ey}. The trade-off between stability and responsivity makes it difficult \cite{Kang.2012} to obtain both stable flow and a fast flow response with a syringe pump system, both of which are necessary for LSCI characterization measurements.

Pressure-driven flow control systems overcome the limitations of syringe pumps by eliminating the use of a motor and lead screw to provide very stable flow \cite{Bong.2010,Zhou:2011ey,Li:2014ca,Zeng.2020}. These systems use pressure-regulated air to push liquid from a reservoir through the microfluidic device at a constant flow rate \cite{Zeng.2020}. In order to control the applied pressure, an absolute flow sensor must be added inline with the microfluidic device. Thermal mass flow sensors are commonly used to measure the low flow rates found in microfluidics. These sensors use heating and temperature-sensing elements to detect the magnitude and direction of flow and are factory-calibrated to measure absolute flow rates for various fluids. Integration of the flow sensor with a feedback loop allows for the programming of specific flow rates rather than pressure levels.

@@

The microfluidic flow phantom was cast using PDMS (Sylgard 184™, Dow Corning) with 1.8 mg titanium dioxide (\ce{TiO2}) added per gram of PDMS to generate a scattering background that mimics tissue optical properties at near infrared wavelengths ($\mu_s' = 8$ cm$^{-1}$) \cite{Parthasarathy:2008el,Yaroslavsky:2002tg}. The PDMS-\ce{TiO2} mixture was vigorously mixed for 15 minutes to ensure even distribution of the \ce{TiO2} in the phantom. Centrifugation (3000 g relative centrifugal force, 10 minutes) was then performed to de-gas the sample and remove any \ce{TiO2} clumps.

A machined aluminum mold was used to cast the microfluidic as shown in Fig.~\ref{fig:microfluidic}a. The extrusion is 460 $\mu$m $\times$ 460 $\mu$m $\times$ 25 mm, which formed three walls of the channel. The square cross-section (1:1 aspect ratio) approximates the circular cross-section of vascular geometries \textit{in vivo}. The PDMS-\ce{TiO2} mixture was carefully poured on top of the mold within a polystyrene petri dish and placed in a vacuum chamber to remove air bubbles introduced during pouring. The PDMS was then cured at 70 $^\circ$C for one hour, allowed to cool, and detached from the mold. The microchannel side of the phantom was then oxygen plasma bonded to a glass slide to seal the channel. Inlet and outlet ports were attached using 23 gauge blunt tip needles fitted partially inside the phantom, which were epoxied in place for an air-tight seal. Tygon® microbore tubing (0.02$''$ inner diameter, Cole-Parmer Instrument Company, LLC) was connected to the blunt tip ports as inlet and outlet tubing. The inlet tubing was connected to the flow control system using fluidic fittings with FEP tubing (IDEX Health and Science, LLC). The outlet tubing drained into a vial, with the tubing end fully submerged in liquid (see Supplementary Material S1). A photograph of the phantom is shown in Fig.~\ref{fig:microfluidic}b, where the channel is visible inside the \ce{TiO2}-doped PDMS with the glass slide sealing the surface.

@@

To assess the performance of the syringe pump and pressure-regulated flow systems, and to quantify the uncertainty of each system, it is necessary to know the true flow rate of the blood-mimicking solution through the microfluidic channel. A thermal mass flow sensor (FLU-L, 0 $\pm$ 1 mL/min, Fluigent, Inc.) was placed inline between the flow control system and the phantom inlet tubing, as shown in Fig.~\ref{fig:system_lsci}a. This flow rate range corresponds to flow speeds of 0 $\pm$ 77.8 mm/s in the microfluidic channel. The sensor output was relayed to a computer via an accompanying hub (Flowboard, FLB, Fluigent, Inc.) where the MAESFLO software (Fluigent, Inc.) was used to monitor and record the absolute flow rate. All flow sensor measurements were downsampled from 10 Hz to 2 Hz using interpolation to match the timing of the optical measurements.

@@

Two separate stepped flow profile experiments were conducted to assess each flow system using the commercial flow sensor. First, a 3-step flow profile was used to determine the uncertainty of each flow control system. The experiment lasted 240 seconds, beginning with 30 seconds of no flow (0 mm/s) followed by one full minute at each flow speed (2.4, 3.6, and 4.8 mm/s) before concluding with an additional 30 seconds of no flow. The experiment was repeated three times on each system. The second flow profile experiment consisted of a 13-step flow profile. The flow speed was once again varied from 2.4 to 4.8 mm/s, however in smaller 0.2 mm/s increments. As with the first experiment, the flow profile begins and ends with 30 seconds of no flow, with one full minute at each programmed flow speed (840 seconds total duration). The second experiment was also repeated three times on each flow control system.

@@

\noindent where $T$ is the camera exposure time, $x=T/\tau_{c}$, and $\beta$ is an instrumentation factor that accounts for
speckle sampling, polarization, and coherence effects \cite{Bandyopadhyay:2005bg}. For these measurements, it was assumed that $\beta=1$, which exclusively allows for the computation of relative flow changes. The resulting timecourse of $\tau_c$ values was resampled to 2 Hz using interpolation and smoothed temporally with a central moving average filter ($k=5$).

@@

A pair of consumer camera lenses were used to image the scattered light (AF Nikkor 50 mm $f$/1.8D + AF Micro-Nikkor 105 mm $f$/2.8D, Nikon Corp.) with 2.1$\times$ magnification through a bandpass filter (785$\pm$31 nm, 87-773, Edmund Optics Inc.) to a CMOS camera (acA1920-155um, 1920 $\times$ 1200 pixels, Basler AG). The aperture of the lower lens was set to $f$/5.6 while the upper lens was left at its maximum size of $f$/2.8. Only a subset of the overall sensor array was used (1000 $\times$ 750 pixels), resulting in a field-of-view of 2.9 $\times$ 2.2 mm. The camera exposures were temporally synchronized with the modulated laser pulses from the AOM. Fifteen different camera exposures ranging between 50 $\mu$s and 80 ms were recorded for each complete MESI frame, resulting in an effective acquisition rate of approximately 2.5 fps. The total amount of light used to capture each exposure was held constant with the AOM in order to minimize the effects of shot noise \cite{Parthasarathy:2008el}. The entire acquisition was controlled using custom software written in C++ integrated with a multifunction I/O device (USB-6363, National Instruments Corp.) for the generation of camera exposure trigger signals and AOM modulation voltages.

@@

A stepped flow profile experiment was conducted using the pressure-regulated flow system to assess the performance of both MESI and LSCI using the same optical hardware. The flow profile consisted of 19 steps spanning 1 -- 10 mm/s in 0.5 mm/s increments. Each speed was maintained for two minutes with a total experiment duration of 40 minutes, including brief periods of no flow (0 mm/s) at the beginning and end of the profile. MESI measurements were acquired continuously throughout the experiment at the maximum acquisition rate of the system ($\sim$2.5 fps). Single-exposure (0.5 ms) LSCI frames were extracted from the MESI dataset \textit{post hoc}. The initial rise of the first flow step was again used to synchronize the timing of the MESI/LSCI and flow sensor datasets. The experiment was repeated a total of four times across two days.

@@

\noindent where $T$ is the camera exposure time, $x=T/\tau_c$, $\rho$ is the fraction of light that is dynamically scattered, $\beta$ is a normalization factor that accounts for speckle sampling, and $\nu$ represents exposure-independent instrument noise and nonergodic variances. $\beta$ was held constant at all timepoints during an experiment by performing an initial fit to Eq. \ref{eq:mesi} using the median speckle contrast value at each exposure across the entire multi-exposure dataset. All fitting was performed with the Levenberg-Marquardt nonlinear least squares algorithm \cite{Lourakis:J2fCMU5i} using a custom program written in MATLAB (R2021a, MathWorks, Inc.). The resulting timecourse of $\tau_c$ values was resampled to 2 Hz using interpolation, smoothed temporally with a central moving average filter ($k=5$), and inverted to obtain $ICT$.

@@

Almost all microfluidic studies used for characterizing LSCI and MESI systems have been performed using syringe pumps. This study corroborates previous work \cite{Li:2014ca,Zhou:2011ey,Zeng.2020} that has shown syringe pumps do not produce stable flow outputs and demonstrates that single-exposure LSCI is sensitive enough to detect the resulting flow anomalies (Fig. \ref{fig:SvP-FS_LSCI}). Furthermore, the close agreement between LSCI and the flow sensor demonstrates that these anomalies could be incorrectly interpreted as noise in the optical imaging system when they are actual flow variations. Such a conclusion would underestimate the true performance of an LSCI or MESI system being evaluated for real-world applications. Compared to the syringe pump, the pressure-regulated flow system produced much more stable and reproducible flow outputs across all trials. Because microfluidic studies are commonly used to assess LSCI and MESI instrument performance, an unstable flow control system contributes additional noise to the measurements, making it more difficult to properly characterize a system. For example, the flow profiles in Fig. \ref{fig:SvP-FS} and the uncertainty analysis in Table \ref{tab:uncertainty_combined} show large errors and high variability within and across trials using the syringe pump. These instabilities in flow were caused entirely by the syringe pump, as independently validated using single-exposure LSCI in Fig. \ref{fig:SvP-FS_LSCI}a. Such instability would increase the uncertainty in the true performance of any instrument being evaluated. The same analysis shows that a more reliable flow controller, such as the pressure-regulated flow system, produces minimal error and much improved reproducibility, therefore facilitating a more robust instrument assessment.

@@

The linearity of the results seen with both MESI and LSCI suggest that a calibration to absolute flow speeds might be feasible. However, any such calibration would be highly dependent upon experimental parameters, including the sample geometry (i.e. channel size), scatterer concentration, illumination configuration, and exposure time selection. Changes to any of these parameters would produce different $ICT$ values, which would affect the correlation with the flow sensor and invalidate the calibration. It also remains to be seen whether calibrations performed in controlled microfluidic environments can reliably be translated to animal or human subjects since \textit{in vivo} measurements sample a range of different vessels \cite{Davis:2016ik,Jafari.2020}. However, if microfluidics are necessary for the calibration process, these results demonstrate that a pressure-regulated flow system should be used in order to produce stable, repeatable, and accurate flow.

@@

\subsection*{Acknowledgments}
The authors would like to acknowledge Jason R. Cook for machining the aluminum mold used to cast the microfluidic channel. The authors would also like to thank Sertan Kutal Gökçe for providing the oxygen plasma bonding procedure, which was used to fabricate the microfluidic phantom. This study was supported by the National Institutes of Health (Nos. NS108484, EB011556, and EB007507) and the UT Austin Portugal Program.
